---
title: "Lab Report 5 and Read Me"
author: "Faith"
date: "`r format(Sys.Date())`"
output: github_document
---
## Welcome

Welcome to my Lab 5 repository and report! In this lab we worked on summarizing data in various ways.

## Lab Objectives

- Rename variables
- Create new variables
- Summarize a dataset
- Summarize a dataset by groups
- Visualize the relationship between variables
  - Numerical and catagorical variable
  - Two numerical variables
  - Three variables

## Lab Prep

Packages were loaded.

```{r}
library(tidyverse)
```

## Reformat the data

```{r}
iris <- as_tibble(iris)
iris
```

## Assignment Questions

1. Rename each variable so that it is all lower-case and uses an underscore _ instead of a period . in the name (the recommended coding style in the tidyverse style guide). Print the resulting table.

```{r}
new_titles <- rename(iris,
       sepal_length = Sepal.Length,
       sepal_width = Sepal.Width,
       petal_length = Petal.Length,
       petal_width = Petal.Width,
       species = Species
       )

new_titles
```

2. Convert the four numerical variables from cm to mm by multiplying by 10. Print the resulting table.

```{r}
mm <- mutate(new_titles, sepal_length = sepal_length * 10,
       sepal_width = sepal_width * 10,
       petal_length = petal_length * 10,
       petal_width = petal_width * 10,
       )

mm
```

3. Calculate sepal area and petal area (area is equal to length multiplied by width). Print a table with only the variables sepal area, petal area, and species.

```{r}
area <- mutate(mm, sepal_area = sepal_length * sepal_width,
       petal_area = petal_length * petal_width
       )

species_area <- select(area, sepal_area, petal_area, species)

species_area
```

4. Calculate the following statistics for the entire dataset from the sepal length variable and print the resulting table:

sample size
maximum value
minimum value
range
median
first quartile (q1)
third quartile (q2)
inter-quartile range (iqr)

```{r}
sepal_summary <- summarize(mm, sampl_size = n(),
        max = max(sepal_length),
        min = min(sepal_length),
        range = max - min,
        median = median(sepal_length),
        q1 = quantile(sepal_length, .25),
        q3 = quantile(sepal_length, .75),
        IQR = IQR(sepal_length)
)

sepal_summary
```

5. Calculate the following statistics for each species from the petal width variable and print the resulting table:

sample size
mean
standard deviation
variance
standard error of the mean
approximate 95% confidence interval

```{r}
species <- group_by(mm, species)

species

new <- summarize(species, sampl_size = n(),
          mean = mean(petal_length),
          sd = sd(petal_length),
          var = var(petal_length),
          se = sd / sqrt(sampl_size),
          UCL = mean + (2*se),
          LCL = mean - (2*se)
)

new

```

6. Visualize the relationship between petal length and species using a strip plot.

```{r}
ggplot(data = species) +
        geom_jitter(mapping = aes(x = species, y = petal_length))
```

7. Starting with the previous graph, add the mean and 95% confidence interval for each species.

```{r}
ggplot(data = species) +
        geom_jitter(mapping = aes(x = species, y = petal_length)) +
        geom_crossbar(
                data = new, 
                mapping = aes(x = species, y = mean, ymax = UCL, ymin = LCL),
                color = "red"
        )
```

8. Visualize the relationship between petal length, petal width, and species using a scatterplot. Map the two numerical variables to the x and y axes and map species to the color and shape aesthetics.

```{r}
ggplot(data = species) +
        geom_point(mapping = aes(x = petal_length, y = petal_width, color = species, shape = species))
```

## Session Info

Here is my Session information:

```{r}
sessioninfo::session_info()
```